---
title: "Plotting ECS distirbutions from estimated gamma distributions"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal 

Produce a figure with the ECS samples. This may serve as a good supplemental figure for the paper to show the distribution shapes for the evidence configurations from S20, based on gamma distribution assumption. 

```{r}
library(tidyverse)
```
# Produce figure

We produce PDF curves in a panel using the color palette and then save the figure in the `figures` directory.

This code will will produce the publication level figure. A simple visualization is available in `02_fitting_andsampling_ecs_distributions.Rmd`.

```{r}
ecs_sample_list <- readRDS("data/ecs_samples_plot_data.RDS")

# convert to dataframe
ecs_df_S1 <- data.frame(
  
  value = unlist(ecs_sample_list),
  evidence = rep(names(ecs_sample_list), each = n),
  row.names = NULL
  
)

# recode evidence configurations for aesthetics
ecs_df_S1$evidence <- ecs_df_S1$evidence %>%
  recode(
    IPCC = "IPCC AR6",
    No_Process = "No Process",
    No_Historical = "No Historical",
    No_Paleoclimate = "No Paleoclimate",
    Baseline_Emergent_constraints = "Baseline + Emergent constraints"
  )

# edit order of the facet panels
facet_order <- c(
  "IPCC AR6",
  "Baseline",
  "No Historical",
  "No Process",
  "No Paleoclimate",
  "Baseline + Emergent constraints"
)

# convert the evidence configuration to factor with the facet order we want
ecs_df_S1$evidence <- factor(ecs_df_S1$evidence, levels = facet_order)

# plot with ggplot"
ggplot() +
  geom_density(data = ecs_df_S1,
               aes(x = value, 
                   color = evidence,
                   fill = evidence),
               linewidth = 0.75, 
               alpha = 0.2, 
               bw = 0.3) +
  scale_color_manual(values = ECS_COLORS, 
                     name = "ECS Configurations") +
  scale_fill_manual(values = ECS_COLORS, 
                    name = "ECS Configurations") +
  scale_x_continuous(breaks = seq(from = 0, to = 9, 
                                  by = 1)) +
  labs(x = "Equilibrium Climate Sensitivity Value",
       y = "Densisty") +
  facet_wrap(~ evidence) +
  theme_light() +
  theme(legend.position = "bottom")

ggsave("figures/S1_fig_ECS_distributions.png",
       device = "png",
       width = 7,
       height = 5,
       units = "in",
       dpi = "print")
```
