---
title: "Producing Figures"
author: "Joe Brown"
date: "2024-06-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal

The goal of this script is to produce figures to visualize results of ECS evidence analysis.

```{r, message=FALSE}
library(matilda)
library(tidyverse)
library(MASS)
library(spatstat)
```

# Weighted ensemble

```{r}
weighted_gsat_ensemble <- readRDS("data/weighted_gsat_ensemble.RDS")

eoc_gsat <- lapply(names(weighted_gsat_ensemble), function(df_name) {
  
  data <- weighted_gsat_ensemble[[df_name]]
  
  data$scenario <- df_name
  
  return(data)
  
})


eoc_gsat_df <- do.call(rbind, eoc_gsat)

ggplot(data = eoc_gsat_df) +
  geom_line(aes(x = year, 
                y = value, 
                group = run_number, 
                color = weight_norm, 
                alpha = weight_norm), 
            linewidth= 0.1) +
  scale_color_gradient(low = "lightblue", high = "dodgerblue4", name = "Weights") +
  scale_alpha_continuous(range(c(0,1))) +
  labs(x = "Years", y = "Temp") +
  theme_light() +
  guides(alpha = "none") +
  facet_wrap(~ scenario)
  
ggsave("figures/S2_fig_gsat_ensemble.png",
       device = "png",
       width = 28,
       height = 15,
       units = "cm",
       dpi = "print")
```

# Median temperature and confidence interval

Compute the weighted median and confidence intervals.

```{r}
# median (CI) warming calculation
warming_projection <- eoc_gsat_df %>% 
  group_by(year, scenario) %>% 
  summarize(
    median = weighted.quantile(x = value, w = weight_norm, probs = 0.5), 
    ci_5 = weighted.quantile(x = value, w = weight_norm, probs = 0.05),
    ci_16 = weighted.quantile(x = value, w = weight_norm, probs = 0.16), 
    ci_84 = weighted.quantile(x = value, w = weight_norm, probs = 0.84), 
    ci_95 = weighted.quantile(x = value, w = weight_norm, probs = 0.95),
    ci_66 = weighted.quantile(x = value, w = weight_norm, probs = 0.66), 
    ci_33 = weighted.quantile(x = value, w = weight_norm, probs = 0.33),
    .groups = "drop")

```

Plotting temperature projections:

```{r}
# recoding scenario names for the plot
warming_projection$scenario <- recode_factor(
  warming_projection$scenario,
  No_Historical = "No Historical",
  No_Process = "No Process",
  No_Paleoclimate = "No Paleoclimate",
  Baseline_Emergent_constraints = "Baseline + Emergent constraints")


# Order factor levels for plot facet 
facet_order <- c("Baseline",
                 "No Process",
                 "No Historical",
                 "No Paleoclimate",
                 "Baseline + Emergent constaints")

```

Construct plot: 

```{r}
temp_projection_plot <- 
  ggplot(data = subset(warming_projection, 
                       year > 2023)) +
  geom_line(aes(x = year, 
                y = median, 
                color = scenario), 
            linewidth = 0.75) +
  scale_color_manual(values = ECS_COLORS) +
  geom_ribbon(aes(x = year, 
                  ymin = ci_5, 
                  ymax = ci_95, 
                  fill = scenario, 
                  color = scenario), 
              alpha = 0.1, 
              linetype = "blank") +
  # geom_ribbon(aes(x = year, 
  #                 ymin = ci_16, 
  #                 ymax = ci_84, 
  #                 fill = scenario, 
  #                 color = scenario), 
  #             alpha = 0.2, 
  #             linetype = "blank") +
  scale_fill_manual(values = ECS_COLORS) +
  labs(x = "Year", 
       y = expression(paste("Future Warming (GSAT) relative to 1995-2014 (", degree, "C)"))) +
  theme_light(base_size = 16) +
  theme(legend.position = "none") +
  facet_wrap(~ scenario)

temp_projection_plot
```

```{r}
historic_tas_subset <- subset(eoc_gsat_df, 
                                year > 1849 &
                                year < 2024)

historic_plot_data <- historic_tas_subset %>% 
  group_by(scenario, year) %>% 
  summarize(
    median = weighted.quantile(value, w = weight_norm, probs = 0.5),
    lower_ci = weighted.quantile(value, w = weight_norm, probs = 0.05),
    upper_ci = weighted.quantile(value, w = weight_norm, probs = 0.95),
    .groups = "drop")

```

```{r}
# recoding scenario names for the plot
historic_plot_data$scenario <- recode_factor(
  historic_plot_data$scenario,
  No_Historical = "No Historical",
  No_Process = "No Process",
  No_Paleoclimate = "No Paleoclimate",
  Baseline_Emergent_constraints = "Baseline + Emergent constraints")


# Order factor levels for plot facet 
facet_order <- c("Baseline",
                 "No Process",
                 "No Historical",
                 "No Paleoclimate",
                 "Baseline + Emergent constaints")

# normalize obs historical temperature
## normalization function
data_normalization <- function(data, reference_years) {
  
  reference_data <- data[data$year %in% reference_years, ]
  
  mean_reference_period <- mean(reference_data$value)
  
  normalize_values <- data$value - mean_reference_period
  
  data$value <- normalize_values
  
  return(data)
  
}


# normalizing historical data to the reference period
reference_historical <- data_normalization(temp_hist, 1995:2014)
```


```{r}
temp_projection_plot +
    geom_point(data = reference_historical, 
             aes(x = year, 
                 y = value), 
             color = "black", 
             size = 0.3) + 
  geom_line(data = historic_plot_data, 
            aes(x = year, 
                y = median), 
            color = "#92397a",
            linewidth = 0.7) +
  geom_ribbon(data = historic_plot_data, 
              aes(x = year, 
                  ymin = lower_ci, 
                  ymax = upper_ci), 
              fill = "#92397a",
              alpha = 0.2)


ggsave("figures/temp_constrained_warming_projections.png",
       device = "png", 
       units = "in",
       width = 10,
       dpi = "print")
```

# Temperature probability

Plotting temperature probabilities for each of the ECS evidence configurations.

Loading data:
```{r}
temp_probability <- readRDS("data/temp_probability_results.RDS")

temp_probability_df <- do.call(rbind, temp_probability)
row.names(temp_probability_df) <- NULL

temp_probability_df <- temp_probability_df %>% 
  complete(scenario, bins) %>% 
  mutate(bins = recode(bins, !!!bins_mapping))

temp_probability_df[is.na(temp_probability_df)] <- 0
```

```{r}
# recoding scenario names for the plot
temp_probability_df$scenario <- recode_factor(
  temp_probability_df$scenario,
  Baseline = "Baseline",
  No_Historical = "No Historical",
  No_Process = "No Process",
  No_Paleoclimate = "No Paleoclimate",
  Baseline_Emergent_constraints = "Baseline + Emergent constraints")

# Order factor levels for plot 
scenario_order <- 
  c("Baseline",
    "No Process",
    "No Historical",
    "No Paleoclimate",
    "Baseline + Emergent constraints")

# order factor levels
temp_probability_df$scenario <- factor(temp_probability_df$scenario, 
                                       levels = scenario_order)
```

The `bang bang bang` operator (`!!!`) expands vectors/lists into individual arguments. Here, this operator allows us to replace the `bins` factor levels according to the vectors in `bins_mapping`. 

Constructing plot:
```{r}
probability_plot <- 
  ggplot(data = temp_probability_df, 
         aes(fill = bins, 
             x = scenario, 
             y = probability)) +
  geom_bar(position = position_fill(reverse = T), 
           stat = "identity", 
           width = 0.6) +
  scale_y_continuous(breaks = seq(0.0, 1.0, 0.1)) +
  scale_fill_manual(
    values = TEMP_PROBABILITY_COLORS,
    labels = bins_mapping,
    name = "Warming") +
  labs(y = "Probability", 
       x = NULL) +
  coord_flip() +
  theme_light(base_size = 16) +
  theme(legend.position = "bottom", 
        axis.text = element_text(size = 14))

probability_plot

```

Save the plot:
```{r}
# save the probability plot
ggsave("figures/fig_NUM_temperature_probability_plot.png", 
       probability_plot, 
       device = "png", 
       width = 12,
       height = 8, 
       units = "in", 
       dpi = "print")

```

# GSAT stats plot

This plot shows the spread, distribution shape and some descriptive statistics of the end-of-century `gsat` metrics.

```{r}
gsat_metrics <- readRDS("data/weighted_gsat_metrics.RDS")

gsat_metrics_df <- do.call(rbind, gsat_metrics)
gsat_metrics_df$scenario <- factor(gsat_metrics_df$scenario)

gsat_metrics_stats <- data_summary(gsat_metrics_df)

```

```{r}
gsat_metrics_df <- recode_scenarios(gsat_metrics_df)
gsat_metrics_stats <- recode_scenarios(gsat_metrics_stats)

```

Constructing plot:
```{r}
gsat_stats_plot <- 
  ggplot() +
    geom_jitter(data = gsat_metrics_df, 
              aes(x = scenario, 
                  y = metric_result,
                  color = scenario),
              stat = "identity",
              width = 0.20,
              alpha = 0.1, 
              size = 0.7) +
  geom_violin(data = gsat_metrics_df, 
              aes(x = scenario, 
                  y = metric_result, 
                  color = scenario),
              linewidth = 0.7,
              alpha = 0.0,
              trim = FALSE) +
  geom_errorbar(data = gsat_metrics_stats, 
                aes(x = scenario, 
                    y = median, 
                    ymin = lower, 
                    ymax = upper, 
                    color = scenario), 
                width = 0.2, 
                linewidth = 0.7) +
  geom_point(data = gsat_metrics_stats,
             aes(x = scenario, 
                 y = median, 
                 color = scenario), 
             size = 6) +
  scale_color_manual(values = ECS_COLORS, name = "ECS configuration") +
  geom_label(data = gsat_metrics_stats, 
             aes(x = scenario, 
                 y = median + 0.2, 
                 label = paste(round(median, 2), "\n[", round(lower, 2), ",", round(upper, 2), "]")), 
             size = 3) +
  labs(y = "Future Warming (GSAT) relative to 1995-2014") +
  theme_light(base_size = 14) +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        legend.position = "inside",
        legend.position.inside = c(0.15, 0.89),
        legend.background = element_rect(fill = "white", colour = "lightgrey", linewidth = 0.5, linetype = "solid"),
        axis.title.x = element_blank(), 
        axis.text = element_text(size = 12), 
        axis.text.x = element_blank())

gsat_stats_plot
```

```{r}
ggsave("figures/fig_NUM_temperature_metric_distribution.png",
       gsat_stats_plot,
       device = "png", 
       width = 8, 
       height = 8, 
       units = "in", 
       dpi = "print")
```


# Constrained and weighted ECS distribution

This figure is meant to demonstrate the distribution of ECS values that resulted in the temperature projections and temperature probabilities.

The data is built by merging the sampled ECS distributions with the constrained and weighted `gsat` metrics. Here we merge both the ECS values and the RMSE values to the . 

```{r}
ecs_samples <- readRDS("data/ecs_samples_plot_data.RDS")

ecs_df <- do.call(rbind, ecs_samples)
row.names(ecs_df) <- NULL
ecs_df$run_number <- rep(seq(1, 15000), times = 5)
ecs_df$scenario <- rep(scenario_order, each = 15000)

ecs_plot_df <- merge(gsat_metrics_df, ecs_df, by = c("run_number", "scenario"))

```

# Construct plot:

```{r}
ecs_dist_plot <- 


```


